Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-09-21 manuscript/manuscript.tex
Edit the delta value determination section
This was an easy one as the text was already very good.
I removed a bit of redundancy and shortened the text a bit.

For an exponential density function, the density-based discretization will be applied
before the adaptive discretization algorithm.
This means that optimal values for both the distance-size ratio $D$ and the delta ratio
$\delta$ must be determined.
We perform an error analysis similar to what was done for the linear density case.
However, we only consider the "Big Grid" scenario from Table \ref{tab:grids} for the
sake of brevity.

@@

Having well controlled values for $D$, we are free to explore the integration error as a
function of $\delta$ in more detail and how it varies for different values of $b$ (Eq.
\ref{eq:density-exp}).

\begin{figure}
\centering
\includegraphics[width=\linewidth]
    {figures/grid-search.pdf}
\caption{
    Numerical error exploration in the $D$-$\delta$ space.
    The percentage difference values were obtained from the comparison
    between the analytical solution and the numerical approximation of
    the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
    spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
    These comparisons were carried out on the "Big Grid" configuration (Table
    \ref{tab:grids}), with a spherical shell of
    35km of thickness, and density function with $b$ of 1km.
    The points inside the dashed line are the ones that present an
    error lower than 0.1\%.
    }
\label{fig:grid-search}
\end{figure}

@@

In order to determine a default value of $\delta$ that ensures an
acceptable accuracy for every exponential density case we have
performed comparisons between the numerical model and the analytical
solutions, keeping the distance-size ratio $D$ equal to the values
obtained for the linear density case while exploring different values of
$\delta$.

Firstly, we have considered the same spherical shell models we have
used in the linear density case: a thin one (1km of thickness) and a
thick one (35km of thickness) composed by $30^\circ \times 30^\circ$
tesseroids.
But in this case, we assigned an exponential density function like the
one shown in equation \ref{eq:density-exp} to each shell, choosing
a few values of $b$ in order to explore different density
variations.

Finally, we have computed the gravitational potential ($V$), the
vertical component of its gradient ($g_z$) and the $g_{zz}$ component of
the Marussi tensor on the big grid (see Table \ref{tab:grids}) and
calculated the percentage difference between these numerical results and
their corresponding analytical solutions.
Our intention is to set the default $\delta$ as the highest value whose
numerical error is bellow the $0.1\%$ threshold.

On Figure \ref{fig:delta-exponential} we show the density functions
used on each comparison and the resulting percentage difference for each
gravitational field.
The first column shows the results for the thin spherical shell, while the
second one shows the ones corresponding to the thick shell.
From these we can conclude that a $\delta = 0.2$ is needed in order to
achieve an acceptable numerical error for the computation of the
potential and the $g_z$ component.
Because the same comparison using the other components of the gradient ($g_x$
and $g_y$) could not be carried out, we will extrapolate this result
setting a default $\delta = 0.2$ for their computation.

On the other hand, the comparison with the $g_{zz}$ component of the
Tensor shows a difference between the thin and the thick spherical
shell: the former needs a $\delta = 0.2$, but the latter requires a
$\delta = 0.01$.
In order to guarantee a good accuracy for every case, we will keep the
more conservative value of $\delta = 0.01$ for the computation of the
Marussi tensor components.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
\caption{
    Density functions used in the $\delta$ ratio determination and the
    differences between the gravitational fields generated by the
    numerical model and the analytical solution for a thin and a thick
    spherical shell of 1km adn 35km of thickness, respectively, both
    with the exponential density defined on equation
    \ref{eq:density-exp}.
    A few values of the constant $b$ were used in order to test the
    numerical model for different density variations.
    The computations were performed on the big grid described in
    Table \ref{tab:grids}, with fixed values of the distance-size ratio
    $D$ obtained for in the linear density case ($D=1, 2, 8$ for the
    potential, gradient components and tensor components, respectively)
    and different values of the $\delta$ ratio.
    If the difference is less than 0.1\%, we consider that the model
    has achieved an acceptable accuracy for the corresponding value of
    $\delta$.
    }
\label{fig:delta-exponential}
\end{figure*}

>>>>

For an exponential density function, the density-based discretization will be applied
before the adaptive discretization algorithm.
This means that optimal values for both the distance-size ratio $D$ and the delta ratio
$\delta$ must be determined.
We perform an error analysis similar to what was done for the linear density case.
However, we only consider the ``Big Grid'' scenario from Table \ref{tab:grids} for the
sake of brevity.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]
    {figures/grid-search.pdf}
\caption{
    Numerical error exploration in the $D$-$\delta$ space.
    The percentage difference values were obtained from the comparison
    between the analytical solution and the numerical approximation of
    the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
    spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
    These comparisons were carried out on the ``Big Grid'' configuration (Table
    \ref{tab:grids}), with a spherical shell of
    35km of thickness, and density function with $b$ of 1km.
    The points inside the dashed line are the ones that present an
    error lower than 0.1\%.
    }
\label{fig:grid-search}
\end{figure}

@@

Having chosen values of $D$ equal to linear density case, we are free to explore the
integration error as a function of $\delta$ in more detail and how it varies for
different values of $b$ (Eq. \ref{eq:density-exp}).
We perform the error calculations for the ``Big Grid'' configuration (Table
\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
The calculations are repeated for different values of $b$ to examine the error variation
with the sharpness of the density function.
Because larger $\delta$ values result in fewer tesseroid divisions,
our intention is to find the highest value of $\delta$ whose numerical error is bellow
the $0.1\%$ threshold.

Fig. \ref{fig:delta-exponential} shows the density functions and the resulting relative
differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
shell and tesseroid model.
The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold for
$\delta = 0.2$ in both the thick and thin shell cases.
On the other hand, optimal values of $\delta$ for $g_{zz}$
are $\delta = 0.2$ for the thin shell and $\delta = 0.01$ for the thick shell.
Once again, we opt for the conservative value of $\delta = 0.01$ and sacrifice
performance for the sake of accuracy.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
\caption{
    Numerical error for different exponential density functions and values of the delta
    ratio.
    The top panels show the exponential density functions for different values of $b$
    (Eq. \ref{eq:density-exp}).
    The error computations were performed on the ``Big Grid'' configuration described in
    Table \ref{tab:grids}, with fixed values of the distance-size ratio
    $D$ obtained for the linear density case ($D=1, 2, 8$ for the
    potential, gradient components, and tensor components, respectively).
    The difference is reported as a percentage of the shell values.
    }
\label{fig:delta-exponential}
\end{figure*}
